ASHA 2023 Listening Effort Data Analysis

### 2023-10-28 ASHA 2023 Analysis/data_analysis.Rmd
exporting roi df
Ran initial analyses for pupil dilation data

@@

pupil_roi <- data %>%
  # For this analysis, we are only analyzing pupil dilation for accurate responses
  dplyr::filter(rep_acc == "accurate") %>%
  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
  dplyr::filter(timebins >= 500) %>%
  dplyr::group_by(subject, trial) %>%
  dplyr::filter(timebins <= max(timebins) - 2000) %>%
  ungroup() %>%
  # For the purpose of modeling, we are centering time at 500 ms (i.e. the beginning ROI time)
  dplyr::mutate(time_c = timebins - 500)

# Creating polynomial variables for model

pupil_roi <- gazer::code_poly(pupil_roi, predictor = "time_c", 
                              poly.order = 3, orthogonal = T, draw.poly = F)

@@

m0 <- glmmTMB(pupil.binned ~ poly1 + poly2 + poly3 + 
                (poly1 + poly2 + poly3|subject) + 
                (poly1 + poly2 +poly3|code), 
              data = pupil_roi)

@@

>>>>

First, we ran an unconditional growth model. I first specified a model with the orthogonal polynomials added as fixed effects and subject (i.e. listener) and code (i.e. phrase) as random effects. I also added the orthogonal polynomials as random slopes. I also used the AR(1) covariance structure to specifiy the random effect structure. I was not able to run the model. Therefore, I attempted to respecify the random effect structure by removing the AR(1) specification. This led to a model with singular fit. In response to this, I decided to remove code (i.e.) from the random effect structure in order to simplify it.

@@

pupil_roi <- data %>%
  # For this analysis, we are only analyzing pupil dilation for accurate responses
  dplyr::filter(rep_acc == "accurate") %>%
  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
  dplyr::filter(timebins >= 500) %>%
  dplyr::group_by(subject, trial) %>%
  dplyr::filter(timebins <= max(timebins) - 2000) %>%
  ungroup() %>%
  # For the purpose of modeling, we are centering time at 500 ms (i.e. the beginning ROI time)
  dplyr::mutate(time_c = timebins - 500) %>%
  # Rescaling pupil size 
  dplyr::group_by(subject, trial) %>%
  dplyr::mutate(norm_pupil = as.numeric(scale(pupil.binned))) %>%
  dplyr::ungroup()

# Polynomials

pupil_roi <- gazer::code_poly(pupil_roi, predictor = "timebins", poly.order = 3, orthogonal = T, draw.poly = F)

# Exporting region of interest df

rio::export(pupil_roi, "Cleaned Data/pupil_analysis_roi.csv")

## I kept running into issues using glmmTMB, so I switched to lme4

library(lme4)
library(lmerTest)

@@

m0 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 +
                (poly1 + poly2 + poly3|subject), 
              data = pupil_roi)

@@

# Model 1

m1 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
                (poly1 + poly2 + poly3|subject), 
              data = pupil_roi)

## Model 2

# Model 2 initially failed to converge, so adding a control statement for bobyqa optimizer

control_bobyqa <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e9))

m2 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
             poly1*speaker +
             (poly1 + poly2 + poly3|subject), 
              data = pupil_roi, control = control_bobyqa)

```
## Model 3

m3 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
             poly1*speaker + poly2*speaker +
             (poly1 + poly2 + poly3|subject), 
              data = pupil_roi, control = control_bobyqa)

```
## Model 4

m4 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
             poly1*speaker + poly2*speaker + poly3*speaker +
             (poly1 + poly2 + poly3|subject), data = pupil_roi, 
           control = control_bobyqa)

summary(m4)

anova(m3, m4)
